Towards a Wide-Coverage Grammar

### 2013-06-24 slides/presentation.tex
New colors for presentation

@@

\begin{frame}
  \begin{block}{Abstract languages}
    Languages of (abstract) terms constrained by the constants they are
    built on and their types. Defined by a \textcolor{red}{signature}
    and a \textcolor{green}{type}.

@@

    $$\synt{aime}\ \synt{Laura}\ (\synt{les}\ \synt{tatous}) :
    \textcolor{green}{S}$$
  \end{block}

  \begin{block}{Object languages}
    Languages of (object) terms which are obtained by replacing the
    constants in an existing (abstract) term using a \textcolor{blue}{lexicon}.

    \begin{columns}[t]
      \begin{column}{0.5\textwidth}
        \begin{align*}
          \mathcal{L}(\synt{tatous}) &= \textcolor{blue}{tatous} \\
          \mathcal{L}(\synt{Laura}) &= \textcolor{blue}{Laura}
        \end{align*}
      \end{column}
      \begin{column}{0.5\textwidth}
        \begin{align*}
          \mathcal{L}(\synt{les}) &= \textcolor{blue}{\lambda^{\circ} x.\ les + x} \\
          \mathcal{L}(\synt{aime}) &= \textcolor{blue}{\lambda^{\circ} x y.\ x + aime + y}
        \end{align*}
      \end{column}
    \end{columns}

@@

  Agreement:
    \begin{align*}
      A_{aime_1} &: NP\_\textcolor{blue}{NUM{=}SG} \limp NP\_\textcolor{blue}{NUM{=}SG} \limp S \\
      A_{aime_2} &: NP\_\textcolor{blue}{NUM{=}SG} \limp NP\_\textcolor{blue}{NUM{=}PL} \limp S
    \end{align*}

  Extraction:
    \begin{align*}
      E_{aime_1} &: NP\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}NO} \\
      E_{aime_2} &: NP\_\textcolor{green}{VAR{=}T} \limp NP\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}ROOT} \\
      E_{aime_3} &: NP\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{green}{VAR{=}T} \limp S\_\textcolor{green}{EXT{=}ROOT}
    \end{align*}

  Combined:
  \begin{small}
  \begin{align*}
      C_{aime_1} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}NO} \\
      C_{aime_2} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}T} \limp NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}ROOT} \\
      C_{aime_3} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}T} \limp S\_\textcolor{green}{EXT{=}ROOT} \\
      C_{aime_4} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{blue}{NUM{=}PL}\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}NO} \\
      C_{aime_5} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}T} \limp NP\_\textcolor{blue}{NUM{=}PL}\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}ROOT} \\
      C_{aime_6} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{blue}{NUM{=}PL}\_\textcolor{green}{VAR{=}T} \limp S\_\textcolor{green}{EXT{=}ROOT}
  \end{align*}

@@

  \begin{columns}[c]
    \begin{column}{0.7\textwidth}
      \begin{itemize}
      \item Every node is a language.
      \item An arrow means that a language is produced from another by
        mapping through a \textcolor{green}{lexicon}.
      \item Two arrows converging on a node mean an
        \textcolor{red}{intersection} of languages.
      \end{itemize}

$$
\mathcal{I}_{\mathcal{G}}(v) = \{t \in \Lambda(\Sigma_v)
\mid\ \vdash_{\Sigma_v} t : S_v\}
$$
$$
\mathcal{E}_{\mathcal{G}}(v) = \mathcal{I}_{\mathcal{G}}(v) \cap
\textcolor{red}{\bigcap_{(u,v) \in E}}
\textcolor{green}{\mathcal{L}_{(u,v)}}(\mathcal{E}_{\mathcal{G}}(u)).
$$
    \end{column}
    \begin{column}{0.3\textwidth}
      \includegraphics[width=\textwidth]{../diagrams/algebraic-perspective.pdf}
    \end{column}
  \end{columns}

@@

  \begin{columns}[c]
    \begin{column}{0.7\textwidth}
      \begin{itemize}
      \item \textcolor{red}{Every node is a term.}
      \item An arrow means that a term is produced from another by
        applying a \textcolor{green}{lexicon}.
      \item Two arrows converging on a node mean that the two (abstract)
        terms yield \textcolor{red}{the same} (object) term.
      \end{itemize}

@@

      $t$ belongs to $\mathcal{P}_{\mathcal{G}}(u)$ iff \\$\exists T :
      \textcolor{red}{V(G) \to \Lambda}$ such that:
      \begin{itemize}
      \item $T_u = t$.
      \item For all $v \in V(G)$, $\vdash_{\Sigma_v} T_v : S_v$.
      \item For all $(v,w) \in E(G)$,
        $\textcolor{green}{\mathcal{L}_{(v,w)}(T_v) = T_w}$.
      \end{itemize}
    \end{column}
    \begin{column}{0.3\textwidth}
      \includegraphics[width=\textwidth]{../diagrams/generative-perspective.pdf}
    \end{column}
  \end{columns}
\end{frame}

>>>>

\usepackage{xcolor}

\definecolor{darkgreen}{rgb}{0., 0.6, 0.}
\definecolor{darkgold}{rgb}{8., 0.6, 0.}

@@

\begin{frame}
  \begin{block}{Abstract languages}
    Languages of (abstract) terms constrained by the constants they are
    built on and their types. Defined by a \textcolor{red}{signature}
    and a \textcolor{darkgreen}{type}.

@@

    $$\synt{aime}\ \synt{Laura}\ (\synt{les}\ \synt{tatous}) :
    \textcolor{darkgreen}{S}$$
  \end{block}

  \begin{block}{Object languages}
    Languages of (object) terms which are obtained by replacing the
    constants in an existing (abstract) term using a \textcolor{darkgold}{lexicon}.

    \begin{columns}[t]
      \begin{column}{0.5\textwidth}
        \begin{align*}
          \mathcal{L}(\synt{tatous}) &= \textcolor{darkgold}{tatous} \\
          \mathcal{L}(\synt{Laura}) &= \textcolor{darkgold}{Laura}
        \end{align*}
      \end{column}
      \begin{column}{0.5\textwidth}
        \begin{align*}
          \mathcal{L}(\synt{les}) &= \textcolor{darkgold}{\lambda^{\circ} x.\ les + x} \\
          \mathcal{L}(\synt{aime}) &= \textcolor{darkgold}{\lambda^{\circ} x y.\ x + aime + y}
        \end{align*}
      \end{column}
    \end{columns}

@@

  Agreement:
    \begin{align*}
      A_{aime_1} &: NP\_\textcolor{darkgold}{NUM{=}SG} \limp NP\_\textcolor{darkgold}{NUM{=}SG} \limp S \\
      A_{aime_2} &: NP\_\textcolor{darkgold}{NUM{=}SG} \limp NP\_\textcolor{darkgold}{NUM{=}PL} \limp S
    \end{align*}

  Extraction:
    \begin{align*}
      E_{aime_1} &: NP\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}NO} \\
      E_{aime_2} &: NP\_\textcolor{darkgreen}{VAR{=}T} \limp NP\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
      E_{aime_3} &: NP\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgreen}{VAR{=}T} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT}
    \end{align*}

  Combined:
  \begin{small}
  \begin{align*}
      C_{aime_1} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}NO} \\
      C_{aime_2} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}T} \limp NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
      C_{aime_3} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}T} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
      C_{aime_4} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}PL}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}NO} \\
      C_{aime_5} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}T} \limp NP\_\textcolor{darkgold}{NUM{=}PL}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
      C_{aime_6} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}PL}\_\textcolor{darkgreen}{VAR{=}T} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT}
  \end{align*}

@@

  \begin{columns}[c]
    \begin{column}{0.7\textwidth}
      \begin{itemize}
      \item Every node is a language.
      \item An arrow means that a language is produced from another by
        mapping through a \textcolor{darkgreen}{lexicon}.
      \item Two arrows converging on a node mean an
        \textcolor{red}{intersection} of languages.
      \end{itemize}

$$
\mathcal{I}_{\mathcal{G}}(v) = \{t \in \Lambda(\Sigma_v)
\mid\ \vdash_{\Sigma_v} t : S_v\}
$$
$$
\mathcal{E}_{\mathcal{G}}(v) = \mathcal{I}_{\mathcal{G}}(v) \cap
\textcolor{red}{\bigcap_{(u,v) \in E}}
\textcolor{darkgreen}{\mathcal{L}_{(u,v)}}(\mathcal{E}_{\mathcal{G}}(u)).
$$
    \end{column}
    \begin{column}{0.3\textwidth}
      \includegraphics[width=\textwidth]{../diagrams/algebraic-perspective.pdf}
    \end{column}
  \end{columns}

@@

  \begin{columns}[c]
    \begin{column}{0.7\textwidth}
      \begin{itemize}
      \item \textcolor{red}{Every node is a term.}
      \item An arrow means that a term is produced from another by
        applying a \textcolor{darkgreen}{lexicon}.
      \item Two arrows converging on a node mean that the two (abstract)
        terms yield \textcolor{red}{the same} (object) term.
      \end{itemize}

@@

      $t$ belongs to $\mathcal{P}_{\mathcal{G}}(u)$ iff \\$\exists T :
      \textcolor{red}{V(G) \to \Lambda}$ such that:
      \begin{itemize}
      \item $T_u = t$.
      \item For all $v \in V(G)$, $\vdash_{\Sigma_v} T_v : S_v$.
      \item For all $(v,w) \in E(G)$,
        $\textcolor{darkgreen}{\mathcal{L}_{(v,w)}(T_v) = T_w}$.
      \end{itemize}
    \end{column}
    \begin{column}{0.3\textwidth}
      \includegraphics[width=\textwidth]{../diagrams/generative-perspective.pdf}
    \end{column}
  \end{columns}
\end{frame}
